A Survey of Range-Only SLAM Techniques and Sensors

### 2018-02-16 Paper.tex
Added brief prediction/measurment to EKF
Added a really brief poorly worded description of the update and measurment phase of EKF
Minor Edits
Edits made to paper as per revision R2.

	\section{Introduction} 
		The Simultaneous Localization and Mapping (SLAM) problem is a common aspect of many autonomous robotics applications. SLAM involves estimating the pose of a mobile robot and building a map of its environment with no a priori information. The SLAM problem is a chicken-and-egg type problem, where the pose of the robot is with respect to a global frame, or map, and building a map requires knowledge of the robot's location. There are many algorithms used to solve this problem, most using a landmark-based approach. The concept of a landmark is identifying a distinguishable feature in the environment through camera or laser rangefinder and estimating the distance and orientation of that landmark relative to the robot. Then, as the robot traverses the environment, it can continue observing that landmark, as well as collecting new landmarks, and using the sensed distances and angles combined with odometry readings to refine its estimation of the robot's pose and map. In most SLAM applications, references to the map is synonymous with the locations of the landmarks.

		In outdoor SLAM applications, such as self-driving cars and autonomous farming equipment, global positioning satellite (GPS) readings are used in addition to the vehicle's odometry. While GPS has low resolution, it proves useful as it provides location information, which can help compensate for the accumulated error from drift in odometric sensors. However, GPS isn't available in all environments, such as indoors and underwater. Instead, wireless signals and sonar beacons are often used in indoors and underwater environments, respectively. While these beacons provide distance estimations to the robotic vehicle, they do not inherently provide location information as with GPS. However, these distances are often still useful as they can be used to compensate for the drift that occurs from measuring a vehicle's displacement through integration of inertial sensor readings. Most SLAM algorithms utilize these range values as a corrective measurement, the same way traditional algorithms use sensed landmarks. However, rather than using range and orientation to landmarks, beacons are only capable of providing range. SLAM algorithms which utilize these beacons are referred to as Range-Only SLAM (RO-SLAM).

@@

		This paper surveys the most popular techniques in RO-SLAM applications. Before discussing RO-SLAM techniques, the different types of range sensors are introduced in Section \ref{range only sensors}, followed by initialization of these beacons in Section \ref{beacon initialization}. Solutions to the RO-SLAM state estimation problem are covered in Section \ref{methods and their implementations}. Finally, the conclusion in Section \ref{conclusion} closes with final thoughts regarding the state of the field of RO-SLAM.

	\section{range only sensors}
	\label{range only sensors}
		There are a variety of range sensors which can be utilized as beacons for RO-SLAM. Beacons refer to any device which involves a signal transmission that can be used for the range estimation without line of sight. The most popular beacons can largely be broken down into two categories: acoustic based and radio frequency based. These beacons serve as landmarks within larger SLAM algorithms. The following section surveys the different types of sensors which can be utilized as range beacons.

@@

		\subsection{Radio Frequency}
			Radio frequency encompass a massive spectrum of devices capable of transmitting or receiving transmission on the radio frequency. There are two methods primarily utilized to determine distance from radio frequency. One is by examining the received signal strength indicator (RSSI) and the other is the by measuring the time of flight of the signal. A system presented by \cite{Padmanabhan2000} processes RF signal strength and utilizes nearest neighbors heuristics to achieve localization based on said signal strength.  \cite{Kurth2003a} describes a method for using very small RF tags as beacons throughout the environment.  As a mobile robot moves throughout an environment it emits a query to which any RF tags that are within the transmission range respond. The robot then estimates the distance to the tag based on the delay. The response from the RF tags include a data tag which contains a unique ID number.

@@

			Wireless Fidelity (wi-fi)  is another choice being explored by man researchers as existing networks are already prevalent in many indoor locations and they are generally low cost \cite{Youssef2005} \cite{Sen2013} \cite{Kotaru2015} \cite{Rea2017}. \cite{Youssef2005} presents a Horus system with the aim of high accuracy and low computation cost for wireless local area network location determining. \cite{Sen2013} proposes a method named CUPID which extracts information from the physical layer to process information allowing it to avoid the impact of multipathing. \cite{Kotaru2015} presents a design and implementation for its SpotFi system which can be deployed on top of existing wi-fi infrastructure. \cite{Rea2017} designed a new method for filtering range measurment specifically designed with the large noise found in indoor WiFi radios in mind.

@@

	\section{Beacon Initialization}
		\label{beacon initialization}
		The core motivation of SLAM algorithms is no a priori knowledge of the operating environment. This includes locations of landmarks, including beacons. Because of this, landmarks are appended to the map as they are discovered in traditional SLAM algorithms, using the robot's estimated pose and the sensed distance and orientation to estimate an initial position of the landmark. The landmark's positions is then refined as the algorithm continues to run.

@@

	\subsubsection{Trilateration}
		\cite{Thrun2002} mentions a basic trilateration technique involving geometric intersection. Rings are drawn at different points of the robots path as shown in Fig[a], where the radius of the rings depend on the strength of the received signals from the beacons. This method of initialization is highly ambiguous as the rings can have more than 1 point of intersecting as shown in Fig[b]  or no  intersecting points as shown in Fig[c]. %more papers needed

	\subsubsection{Particle filter}
		This takes a more probabilistic approach to estimating the initial location of the beacons. As mentioned in\cite{Thrun2002a} initially the entire space is represented by particles. These particles are resampled based on importance weights every iteration. There are several resampling methods mentioned in \cite{Li2015}. Eventually the particle set converges to the accurate location of the beacons. Though this method of localization can be computationally expensive it is preferred whenever the noise involved is non Gaussian.

	\subsection{Undelayed/Online initializations}
		In this case the robot simultaneously initializes and localizes itself and the beacons as it moves
		as mentioned in \cite{Caballero2010}. A Gaussian Mixture Model is integrated into the EKF to create a multiple hypothesis filter.  This multiple hypothesis approach is used to solve the RO-SLAM problem and its use allows for the un-delayed initialization of landmarks, however a large number of hypothesis can lead to increased computational burden. \cite{Geneve2015} also puts forward a method that utilizes a Gaussian mixtures for beacon initialization in EKF, but in this method only two hypotheses and a Cartesian plane. \cite{Ahmad2011a} proposes to reduce the computational burden of methods, such as the one proposed by \cite{Caballero2010}, by introducing a novel state vector that eliminates the need for multiple hypothesis landmark representation. The paper proposes that under different circumstances the landmark could be represented using different parameters. Four circumstances are presented in the paper and the method is tested in simulation using both the EKF and a least squares optimization approach. The results showed that when the circumstances can be clearly distinguished between, an increase in performance is seen.

@@

	The Extended Kalman Filter (EKF) is one of the most popular and widespread methods used to solve the SLAM problem. It is used to overcome the assumption of linear state transitions and measurements, which are rarely seen in practical environments \cite{Thrun2002}. The derivation of the EKF is very well documented, as such, detailed explanations can be viewed from a variety of sources such as: \cite{Thrun2002}, \cite{Ribeiro2004}, and \cite{Haykin2001}. When used to solve SLAM the map applied to the EKF is a feature based map, meaning it is composed of observable features (landmarks) which can be distinguished between during re-observation \cite{Thrun2002}. This distinction becomes important when examining range only SLAM (RO-SLAM) where only the range to a landmark, or the range between landmarks, is known. This restriction means that landmarks often need to be manufactured, such as the beacons discussed in previous sections, and physically placed in an environment. However, that does not mean that the exact location of the landmarks, or even there initial existence, is recognized. Initialization is a critical element in many algorithms that make use of the EKF. If the initialization is inaccurate then the convergence to a correct estimation becomes much more difficult for the filter. The EKF has seen a wide variety of implementation ranging from ground based robotics\cite{Djugash2008}, to unmanned air vehicles \cite{Fabresse2016}, and to autonomous underwater vehicles \cite{Olson2006}. This section will proceed to give a summary of the various ways EKF has been implemented in regards to RO SLAM, as well as examine improvements to the general RO SLAM problem that are applied via methods with EKF mind.

@@

	An extension to EKF called relative-over parametrized (ROP) EKF is presented by \cite{Djugash2008}. This extension uses specific parameterization to  better represent the range only data utilized in RO-SLAM. ROP EKF operates in polar coordinates as opposed to the Cartesian space that typical EKF operates. The method provided improved results over EKF when sparse range measurement or data association errors were present. \cite{Herranz2014} presents a comparison of the ROP-EKF method presented by \cite{Djugash2008}, to a smoothing and mapping method presented by \cite{Dellaert2006}. Experimentation conducted by \cite{Herranz2014} on and indoor and outdoor environment show an improvement in the results of the SAM method over ROP EKF.

@@

	A method proposed by \cite{Dios2015} is geared towards improving map building speed and accuracy for UAS dealing with the 3D RO-SLAM problem. The 3D RO-SLAM problem refers to the added complexity that devices such as submersibles \cite{Newman} and aerial vehicles experience when having to operate in an environment where three dimensions must be considered as opposed to the planar environment many robots are assumed to operate in. The method utilizes inter-beacon measurements and a novel mode selection which is used to select a measurement gathering method based on current conditions \cite{Dios2015}. The first mode is noted as mapping (MS) and the second localization (LS). When in mapping mode the measurement are integrated with a high probability to quickly create a map, when in localization mode the probability is lowered to create higher accuracy localizations. Based on experimentation performed by \cite{Dios2015} the method performed well and a general increase in accuracy and map speed was noted, however it is stated that a wider variety of experimentation should be performed to validate the overall robustness and effectiveness of the method.

@@

	It is explored by \cite{Kehagias2006} how interpolation can be used to generate range data that is equally spaced over time as opposed to the irregular time intervals that range data is typically received at and then applied to solve the RO SLAM problem. The method is tested on the EKF and an batch optimization algorithm with a simulated and physical robot. The results provided by the paper show that the interpolation based algorithms gave better landmark estimation than those without, with the exception of the EKF in two datasets where EKF was already performing poorly.

@@

\subsection{Sparse Extended Information Filters}
Need to add references

Information filters work on same underlying Gaussian belief principle as that of EKF but vary in the way the Gaussian beliefs are represented. An IF algorithm is mentioned in\cite{Torres-Gonzalez2014}. In EKF the gaussians are represented in the form of their moments; mean and variance, whereas in a information filter the Gaussians in their conical parameterization which comprises of a information matrix and information vector.Hence what is computational complex in EKF is simple in the other and vice versa. Hence they are dual to each other.
Sparse Extended information filters is derived from the  information filter but it maintains a sparse representation of the information matrix. This makes the SEIF more scalable and efficient as show in\cite{Torres-Gonzalez2014}

The information filter stores the relation between any 2 state elements in the information matrix also known as links. The higher the value of the link the stronger is the relationship between the 2 state variables and the more it is gonna effect the robots state. SEIF maintains a sparse information matrix by eliminating weak inks that are below a certain threshold thus reducing the computational complexity involved .

In \cite{Torres-Gonzalez2014} they further reduce the computational complexity of by utilizing the computational capabilities of the nodes or radio  beacons. Here the robot computes the prediction stage of the iteration, send the information to the beacons where they compute the update of the process. Hence the computer burden on the robot is reduced. In addition, they factored in inter beacon measurements to reduce uncertainty even further. In \cite{Torres-Gonzalez2017} it evaluates and compares  the SEIF method for 3d SLAM using an octorotor In conclusion this method had better estimates and lesser computational burden than the traditional RO SLAM method using EKF.

\subsection{Particle Filters}      %change the references
Particles filters are made ue up various monte carlo algorithms. they are used to estimate states in partially observable Markov chains as shown in \cite{Doucet2001}. In  initial applications particle filters were used in the field of localization where it was used to estimate a robots pose from sensor data. It is considered better and more robust to conventional estimation methods because it was able to solve the global localization[2] and the kidnapped robot problem\cite{Engelson1992} which EKF in comparison has it drawbacks. It has also been at the core higher dimensional computational problems.  
%Description
Initially robot assumes global uncertainty through a set of particles that are distributed all over the map. Hence both the robot and the landmarks are uncertain.As the robot moves around and measures sensor data it resamples the particle set weighted according to the probability of the particle being the right one. There are several resampling methods depending on computational complexities \cite{Li2015} After some iterations only the particles which have a high degree of probability  remain and eventually converge to the state of the robot and the landmarks.
\begin{figure}[h!]
	\centering
	\includegraphics[height=40mm,width=90mm]{Particle_filter_method.JPG}
	\caption{The leftmost image describes the state of the robot when it just initializes the particle filter. The particles are uniformly distributed.The image in the center describes the particle filter after a couple of iterations. Only the particles which have a higher probability are re sampled. The image on the right shows the particle filter eventually converging to estimate the robot location. }

>>>>

	\section{Introduction} 
		The Simultaneous Localization and Mapping (SLAM) problem is a common aspect of many autonomous robotics applications. SLAM involves estimating the pose of a mobile robot and building a map of its environment with no a priori information. The SLAM problem is a chicken-and-egg type problem, where the pose of the robot is with respect to a global frame, or map, and building a map requires knowledge of the robot's location. There are many algorithms used to solve this problem, most using a landmark-based approach. Landmarks are distinguishable feature in the environment which can be observed through devices such as a camera or a laser rangefinder. Said devices are used to estimate the distance and orientation of the landmark relative to the robot. As the robot traverses the environment it continues to observe and collect landmarks. The sensed distance and angle from each landmark is combined with odometry readings to get an refined estimation of the robot pose and map. 
		%It can continue observing that landmark, as well as collecting new landmarks, and using the sensed distances and angles combined with odometry readings to refine its estimation of the robot's pose and map. 
		In most SLAM applications, references to the map is synonymous with the locations and features of the landmarks.

		In outdoor SLAM applications, such as self-driving cars and autonomous farming equipment, global positioning satellite (GPS) readings are used in addition to the vehicle's odometry. While GPS has low resolution, it proves useful as it provides location information, which can help compensate for the accumulated error from drift in odometric sensors. However, GPS isn't available in all environments, such as indoors and underwater. Instead, wireless signals and sonar beacons are often used in indoors and underwater environments, respectively. These beacons provide distance estimations to the robotic vehicle, but do not inherently provide location information as with GPS. However, these distances are often still useful as they can be used to compensate for the drift that occurs from measuring a vehicle's displacement through integration of inertial sensor readings. Most SLAM algorithms utilize these range values as a corrective measurement, the same way traditional algorithms use sensed landmarks. However, rather than using range and orientation to landmarks, beacons are only capable of providing range. SLAM algorithms which utilize these beacons are referred to as Range-Only SLAM (RO-SLAM).

@@

		This paper surveys state of the art RO-SLAM applications. Before discussing RO-SLAM techniques, the different types of range sensors are introduced in Section \ref{range only sensors}, followed by initialization of these beacons in Section \ref{beacon initialization}. Solutions to the RO-SLAM state estimation problem are covered in Section \ref{methods and their implementations}. Finally, the conclusion in Section \ref{conclusion} closes with final thoughts regarding the state of the field of RO-SLAM.

	\section{range only sensors}
	\label{range only sensors}
		There are a variety of range sensors which can be utilized for RO-SLAM. These range sensors refer to any device which involves a signal transmission that can be used for the range estimation without line of sight. The most popular beacons can largely be broken down into two categories: acoustic based and radio frequency based. These beacons serve as landmarks within larger SLAM algorithms. The following section surveys the different types of sensors which can be utilized as range beacons.

@@

		\subsection{Radio Frequency}
			Radio frequency encompass a massive spectrum of devices capable of transmitting or receiving transmission on the electromagnetic frequency. There are two methods primarily utilized to determine distance from radio frequency. One is by examining the received signal strength indicator (RSSI) and the other is the by measuring the time of flight of the signal. A system presented by \cite{Padmanabhan2000} processes RF signal strength and utilizes nearest neighbors heuristics to achieve localization based on said signal strength.  \cite{Kurth2003a} describes a method for using very small RF tags as beacons throughout the environment.  As a mobile robot moves throughout an environment it emits a query to which any RF tags that are within the transmission range respond. The robot then estimates the distance to the tag based on the delay. The response from the RF tags include a data tag which contains a unique ID number.

			Wireless Fidelity (wi-fi)  is the most popular beacon choice as existing networks are already prevalent in many indoor locations and they are generally low cost \cite{Youssef2005} \cite{Sen2013} \cite{Kotaru2015} \cite{Rea2017}. \cite{Youssef2005} presents a Horus system with the aim of high accuracy and low computation cost for wireless local area network location determining. \cite{Sen2013} proposes a method named CUPID which extracts information from the physical layer to process information allowing it to avoid the impact of multipathing. \cite{Kotaru2015} presents a design and implementation for its SpotFi system which can be deployed on top of existing wi-fi infrastructure. \cite{Rea2017} designed a new method for filtering range measurment specifically designed with the large noise found in indoor WiFi radios in mind.

@@

@@

	\section{Beacon Initialization}
		\label{beacon initialization}
		The core motivation of SLAM algorithms is the exclusion of a priori knowledge of the operating environment. This includes locations of landmarks. Because of this, landmarks are appended to the map as they are discovered in traditional SLAM algorithms, using the robot's estimated pose and the sensed distance and orientation to estimate an initial position of the landmark. The landmark's positions is then refined as the algorithm continues to run.

@@

	\subsubsection{Trilateration}
		Trilateration refers to the process of using the distance to at least three devices at known locations in order to estimate the location of a fourth device. Methods generally utilize the geometry of circles and triangles in order to accomplish this. \cite{Thrun2002} mentions a basic trilateration technique involving geometric intersection. Rings are drawn at different points of the robots path as shown in Fig[a], where the radius of the rings depend on the strength of the received signals from the beacons. This method of initialization is highly ambiguous as the rings can have more than 1 point of intersecting as shown in Fig[b]  or no  intersecting points as shown in Fig[c]. %more papers needed

	\subsubsection{Particle filter}
		Particle filters takes a probabilistic approach to estimating the initial location of the beacons. As mentioned in \cite{Thrun2002a}, initially the entire space is represented by particles. Particles refer to collection of sample states. These particles are resampled based on importance weights every iteration. There are several resampling methods mentioned in \cite{Li2015}. Eventually the particle set converges to the accurate location of the beacons. Although this method of localization can be computationally expensive it is preferred whenever the noise involved is non Gaussian.

	\subsection{Undelayed/Online initializations}
		With undelayed/online initializations simultaneously initializes and localizes itself and the beacons as it moves, as mentioned in \cite{Caballero2010}. A Gaussian Mixture Model is integrated into the EKF to create a multiple hypothesis filter.  This multiple hypothesis approach is used to solve the RO-SLAM problem and its use allows for the un-delayed initialization of landmarks, however a large number of hypothesis can lead to increased computational burden. \cite{Geneve2015} presents a method that models a Gaussian mixtures for beacon initialization in EKF. This method only uses two hypotheses and a Cartesian plane. \cite{Ahmad2011a} proposes to reduce the computational burden of methods, such as the one proposed by \cite{Caballero2010}, by introducing a novel state vector that eliminates the need for multiple hypothesis landmark representation. The paper proposes that under different circumstances the landmark could be represented using different parameters. This representation would lead to reduced computation cost for large scale range only SLAM as opposed to the high computation cost of multiple hypothesis landmark representations. Four circumstances are presented in the paper and the method is tested in simulation using both the EKF and a least squares optimization approach. The results showed that when the circumstances can be clearly distinguished between, an increase in performance is seen.

@@

	The Extended Kalman Filter (EKF) is one of the most popular and widespread methods used to solve the SLAM problem. It is used to overcome the assumption of linear state transitions and measurements, which are rarely seen in practical environments \cite{Thrun2002}. The derivation of the EKF is very well documented, as such, detailed explanations can be viewed from a variety of sources such as: \cite{Thrun2002}, \cite{Ribeiro2004}, and \cite{Haykin2001}. Generally, EKF can be broken up into two phases, the prediction and the measurement phase. The prediction phase updates the state vector based on control inputs and additive Gaussian noise. The measurement phase updates the state vector based observed landmarks with additive Gaussian noise.
	%NOTE THIS PART ABOUT PREDICTION AND MEASURMENT PHASE IS SUPER ROUGH

	 When used to solve SLAM the map applied to the EKF is a feature based map, meaning it is composed of observable features (landmarks) which can be distinguished between during re-observation \cite{Thrun2002}. This distinction becomes important when examining range only SLAM where only the range to a landmark, or the range between landmarks, is known. 
	 %This restriction means that landmarks often need to be manufactured 
	 %add rererence here...
	 %, such as the beacons discussed in previous sections, and physically placed in an environment. 
	 In addition, the exact location of the landmarks, or even there initial existence, is not always known at startup. Initialization is a critical element in many algorithms that make use of the EKF. If the initialization is inaccurate then the convergence to a correct estimation becomes much more difficult for the filter. The EKF has seen a wide variety of implementation ranging from ground based robotics\cite{Djugash2008}, to unmanned air vehicles \cite{Fabresse2016}, and to autonomous underwater vehicles \cite{Olson2006}. This section will proceed to give a summary of the various ways EKF has been implemented in regards to RO SLAM, as well as examine improvements to the general RO SLAM problem that are applied via methods with EKF mind.

@@

	An extension to EKF called relative-over parametrized (ROP) EKF is presented by \cite{Djugash2008}. This extension uses specific parameterization to  better represent the range only data utilized in RO-SLAM. ROP EKF operates in polar coordinates as opposed to the Cartesian space that typical EKF operates. The method provided improved results over EKF when sparse range measurement or data association errors were present. \cite{Herranz2014} presents a comparison of the ROP-EKF method presented by \cite{Djugash2008}, to a smoothing and mapping method presented by \cite{Dellaert2006}. Experimentation conducted by \cite{Herranz2014} on and indoor and outdoor environment show an improvement in the results of the smoothing and mapping (SAM) method over ROP EKF.

@@

	A method proposed by \cite{Dios2015} is geared towards improving map building speed and accuracy for unmanned aerial systems dealing with the 3D RO-SLAM problem. The 3D RO-SLAM problem refers to the added complexity that devices such as submersibles \cite{Newman} and aerial vehicles experience when having to operate in an environment where three dimensions must be considered as opposed to the planar environment many robots are assumed to operate in. The method utilizes inter-beacon measurements and a novel mode selection which is used to select a measurement gathering method based on current conditions \cite{Dios2015}. The first mode is noted as mapping (MS) and the second localization (LS). When in mapping mode the measurement are integrated with a high probability to quickly create a map, when in localization mode the probability is lowered to create higher accuracy localizations. Based on experimentation performed by \cite{Dios2015} the method performed well and a general increase in accuracy and map speed was noted, however it is stated that a wider variety of experimentation should be performed to validate the overall robustness and effectiveness of the method.

@@

	\cite{Kehagias2006} explores how interpolation can be used to generate range data that is equally spaced over time as opposed to the irregular time intervals that range data is typically received at and then applied to solve the RO SLAM problem. The method is tested on the EKF and an batch optimization algorithm with a simulated and physical robot. The results provided by the paper show that the interpolation based algorithms gave better landmark estimation than those without, with the exception of the EKF in two datasets where EKF was already performing poorly.

@@

\subsection{Sparse Extended Information Filters}

Information filters work on same underlying Gaussian belief principle as that of EKF but vary in the way the Gaussian beliefs are represented. An information filter (IF) algorithm is mentioned in\cite{Torres-Gonzalez2014}. In EKF the Gaussians are represented in the form of their moments, mean and variance, whereas in a information filter the Gaussians are in their conical parameterization which comprises of an information matrix and information vector. 
%Hence what is computationally complex in EKF is simple in the other and vice versa. Hence they are dual to each other.
Sparse Extended information filters are derived from the  information filter but it maintains a sparse representation of the information matrix. This makes the SEIF more scalable and efficient as show in \cite{Torres-Gonzalez2014}.

The information filter stores the relation between any two state elements in the information matrix also known as links. The higher the value of the link the stronger is the relationship between the two state variables and the more it is gonna effect the robots state. SEIF maintains a sparse information matrix by eliminating weak inks that are below a certain threshold thus reducing the computational complexity involved.

In \cite{Torres-Gonzalez2014} the computational complexity is reduced by utilizing the computational capabilities of the nodes or radio  beacons. Here the robot computes the prediction stage of the iteration, sends the information to the beacons where the update process is completed. Hence the computation burden on the robot is reduced. In addition, inter-beacon measurements are factored in to reduce uncertainty even further. In \cite{Torres-Gonzalez2017} the SEIF method is evaluated and compared for 3d SLAM using an octorotor In conclusion this method had better estimates and lesser computational burden than the traditional RO SLAM method using EKF.

\subsection{Particle Filters}      %change the references
Particles filters are a category of monte carlo algorithms. they are used to estimate states in partially observable Markov chains as shown in \cite{Doucet2001}. In  initial applications particle filters were used in the field of localization where it was used to estimate a robots pose from sensor data. It is considered better and more robust to conventional estimation methods because of its ability to solve the global localization (THIS IS A FIXED CITATION, PUT ACTUAL CITATION HERE)[2] and the kidnapped robot problem \cite{Engelson1992} which EKF ,in comparison, struggles with. At is core, EKF also has higher dimensional computation problems.
% It has also been at the core higher dimensional computational problems.  
%Description
Initially robot assumes global uncertainty through a set of particles that are distributed all over the map. Hence both the robot and the landmarks are uncertain.As the robot moves around and measures sensor data it resamples the particle set weighted according to the probability of the particle being the right one. There are several resampling methods depending on computational complexities \cite{Li2015} After some iterations only the particles which have a high degree of probability  remain and eventually converge to the state of the robot and the landmarks.
\begin{figure}[h!]
	\centering
	\includegraphics[height=50mm,width=95mm]{Particle_filter_method.JPG}
	\caption{The leftmost image describes the state of the robot when it just initializes the particle filter. The particles are uniformly distributed.The image in the center describes the particle filter after a couple of iterations. Only the particles which have a higher probability are re sampled. The image on the right shows the particle filter eventually converging to estimate the robot location. }
